\documentclass[10pt]{article}
\usepackage{fullpage}
\usepackage{url}
\usepackage{color}
\usepackage{listings}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkblue}{rgb}{0,0,0.7}

\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows}
\tikzset{%
  node distance=2.5cm,  
  initial text={},      
  every state/.style={  
    semithick},
  double distance=2pt,  % Accept state appearance
  every edge/.style={   
    draw,
    ->,
    >=stealth',
    auto,
    semithick} %
}

\lstdefinestyle{jvm}{
  % aboveskip=3mm,
  % belowskip=3mm,
  xleftmargin=2em,
  % xrightmargin=2em,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\ttfamily},
  numbers=left,
  moredelim=[s][\color{black}]{Ljava}{;},
  morecomment=[l][\color{dkgreen}]{;},
  morecomment=[l][\color{magenta}]{;;},
  keywords={class,public,static,super,method,code,end},
  keywordstyle=\color{dkblue},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=8
}

\renewcommand{\thepage}{~}

\title{COM S 440/540 Homework 4}
\date{}
\author{LL Parsing}

\begin{document}

\maketitle

\noindent
Reminder: present your own work and properly cite any sources used.
Solutions should be written satisfying the \emph{other student viewpoint},
and must be prepared using \LaTeX.
\renewcommand{\thepage}{~}
%============================================================
\section*{Question~1~\hfill 30 points}
%============================================================

Consider the grammar below,
for declaring one or more variables in C,
with multiple levels of pointers allowed.
Symbols \emph{type}, \emph{ident}, \verb|*|, \verb|;|, and \verb|,|
are terminals.
\begin{eqnarray}
	S & \rightarrow & \mathit{type} ~ V ~ \verb|;|
\\
	V & \rightarrow & I ~ L
\\
	L & \rightarrow & \verb|,| ~ I ~ L
\\
	L & \rightarrow & \epsilon
\\
	I & \rightarrow & \verb|*| ~ I
\\
	I & \rightarrow & \mathit{ident}
\end{eqnarray}
Build the LL(1) parse table $M$ for this grammar.
For full credit, show how each entry was obtained.

%============================================================
\section*{Question~2~\hfill 20 points}
%============================================================

Based on your answer to the previous question, 
write a predictive, recursive-descent parser,
as pseudocode for each non-terminal grammar symbol.


%============================================================
\section*{Question~3~~(Extra credit for students in 440)\hfill 10 points}
%============================================================

Give a modified version of the grammar
(you may introduce new grammar symbols if needed)
that remains in $LL(1)$,
and the recursive-descent pseudocode that matches it,
to allow for optional \emph{const} designators
(assume \emph{const} is a terminal symbol in the grammar)
according to the following rules.
\begin{enumerate}
	\item
	The initial \emph{type} can be modified by
	a \emph{const} keyword, either before or after
	the \emph{type}, but not both.

	\item
	A \emph{const} keyword may appear after each \verb|*|.
\end{enumerate}
For example, the new grammar should accept
the following sentences (where {\tt int} is a \emph{type}).
\begin{verbatim}
	const int a, b, c;
	int const d, *e, *const f;
	int g, * const h, **const***const*i;
\end{verbatim}

\end{document}
