\documentclass[10pt]{article}
\usepackage{fullpage}
\usepackage{url}
\usepackage{color}
\usepackage{listings}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkblue}{rgb}{0,0,0.7}

\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows}
\tikzset{%
  node distance=2.5cm,  
  initial text={},      
  every state/.style={  
    semithick},
  double distance=2pt,  % Accept state appearance
  every edge/.style={   
    draw,
    ->,
    >=stealth',
    auto,
    semithick} %
}

\lstdefinestyle{jvm}{
  % aboveskip=3mm,
  % belowskip=3mm,
  xleftmargin=2em,
  % xrightmargin=2em,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\ttfamily},
  numbers=left,
  moredelim=[s][\color{black}]{Ljava}{;},
  morecomment=[l][\color{dkgreen}]{;},
  morecomment=[l][\color{magenta}]{;;},
  keywords={class,public,static,super,method,code,end},
  keywordstyle=\color{dkblue},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=8
}

\renewcommand{\thepage}{~}

\title{COM S 440/540 Homework 1}
\date{}
\author{Lexical Analysis 1}

\begin{document}

\maketitle

\noindent
Reminder: present your own work and properly cite any sources used.
Solutions should be written satisfying the \emph{other student viewpoint},
and must be prepared using \LaTeX.
\renewcommand{\thepage}{~}
 
%============================================================
\section*{Question~1~\hfill 15 points}
%============================================================

Write a regular expression
that accepts an HTML special character of the form
\verb|&charname;|
where a \verb|charname| consists
of one or more letters (any case) or digits.
The following are examples of words that should be accepted:
\begin{quote}
\begin{verbatim}
		&frac34;
		&nbsp;
		&circledR;
		&CenterDot;
\end{verbatim}
\end{quote}
You do not need to check that the character name is valid.

%============================================================
\section*{Question~2~\hfill 25 points}
%============================================================

Write a C function, 
with appropriate comments to explain how your code works,
to scan a null-terminated string and return a pointer to the first
``non-whitespace'' character in the string.
If the string is entirely whitespace,
then return a pointer to the terminating null character.
For this exercise, whitespace includes spaces, tabs (\verb|'\t'|),
newlines (\verb|'\n'|), and shell-style comments (between \verb|"#"| and newline).
The function prototype should be
\begin{verbatim}
    const char* skipSpaces(const char* buffer);
\end{verbatim}
and as an illustrating example,
should return the following pointers $p$ for the buffers shown below:
\begin{quote} \tt
\begin{tabbing}
  "\=\verb|\n\t\n|~~\=4"\\
  \> $\uparrow$ \> $\uparrow$ \\
  \> buffer \> $p$ 
\end{tabbing}
\begin{tabbing}
  "\=\# All whitespace \verb|\n|~~~~"\=\\
  \> $\uparrow$ \> $\uparrow$ \\
  \> buffer \> $p$ 
\end{tabbing}
\begin{tabbing}
  "\=~~\verb|\t| \# This is a comment \verb|\n| ~ \# Another \# one \verb|\n\n| ~~\=here\verb|\n\t % #|"\\
  \> $\uparrow$ \> $\uparrow$ \\
  \> buffer \> $p$ 
\end{tabbing}
\end{quote}
You may not call \emph{any} other functions,
except for helper functions you provide.
You will lose points for inefficiency 
(e.g., scanning the string more than once).

\end{document}
